Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalogue Data

### 2012-11-16 doc/acknowledgments.tex
replace accidentally deleted files

TEC acknowledges support from STFC in the form of a research studentship.
%
PJM was given support by the Kavli Foundation and the Royal 
Society, in the form of research fellowships.
%
MWA \ldots
%
SH \ldots
%
SHS \ldots
%
TT acknowledges support from the NSF through CAREER award NSF-0642621,
and from the Packard Foundation through a Packard Research Fellowship.
% 
LVEK acknowledges the support by an NWO-VIDI programme subsidy
(programme number 639.042.505).
%
This research is supported in part by the National Science Foundation under
Grant No. PHY99-07949.

>>>>

TEC acknowledges support from STFC in the form of a research studentship.
%
PJM was given support by the Kavli Foundation and the Royal 
Society, in the form of research fellowships.
%
MWA \ldots
%
SH \ldots
%
SHS, ZG  and TT gratefully acknowledge support from the Packard Foundation in the form of a Packard Research Fellowship to TT and from the National Science Foundation grant AST-0642621.
% 
LVEK acknowledges the support by an NWO-VIDI programme subsidy
(programme number 639.042.505).
%
This research is supported in part by the National Science Foundation under
Grant No. PHY99-07949.

### 2012-11-16 doc/paper-lcr.tex
replace accidentally deleted files

\title[Line of Sight Mass Reconstruction]
{Reconstructing the Lensing Mass in the Universe \\
from Photometric Catalog Data}

\author[Collett \etal]{%
  Thomas~E.~Collett$^{1}$\thanks{\collettemail},
  Philip~J.~Marshall$^{2}$,
  Matthew~W.~Auger$^{1}$,
  Stefan~Hilbert$^{3}$,
\newauthor{%
  Sherry~H.~Suyu$^{4}$,
  Zachary~Greene$^{4}$,
  Tommaso~Treu$^{4}$\thanks{\packard},
  Vasiliy~Belokurov$^{1}$,}
\newauthor{%
  Christopher~D.~Fassnacht$^{5}$,
  L\`eon~V.~E.~Koopmans$^{6}$,
  Roger~D.~Blandford$^{3}$} 
  \medskip\\
  $^1$\ioa\\
  $^2$\oxford\\
  $^3$\kipac\\
  $^4$\ucsb\\
  $^5$\davis\\
  $^6$\kapteyn
}

@@

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight.
Such massive structures contribute an ``external convergence'' $\kappa$ that
in some cases dominates the uncertainty in the inferred distances.  We
characterise this uncertainty by marginalising over a probability density
function for $\kappa$: here, we investigate the use of a simple halo model to
estimate $\pr(\kappa|\data)$ given noisy estimates of the photometric redshift
and stellar mass of galaxies observed in optical imaging surveys of the field
in question. We use mock catalogues from the Millennium Simulation, and compare
our reconstructed $\pr(\kappa|\data)$ to ray-traced $\kappa$ truth values: 
taking into account realistic uncertainties on redshift and stellar masses, we
typically find that the reconstruction yields an improvement in precision of
40\% over the global $\pr(\kappa)$. For any given line of sight the
convergence estimate can be made before the investment of expensive follow-up
observations: we find that selecting systems by the precision with which
$\kappa$ can be estimated leads to a low convergence  sample that would give
time delay distance estimates a factor of 2 more precise than a random
selection,  with biases as low as 0.3\%. Preferentially selecting lines of
sight with high external shear (as might happen in an assembled sample of
quadruply-imaged time delay lenses) has the opposite effect, with distances
biased by around 1\%. We suggest areas for the improvement of this general
analysis framework that should allow yet more accurate cosmological inferences
to be made.

@@

Large numerical simulations have been used to
predict, from first principles, the distribution of likely external
shears and convergences along various lines of sight, including those
containing strong lens systems.  \citet{Holder+Schechter2003} and
\citet{DalalEtal2005} carried out ray-tracing calculations in N-body
simulations to estimate the distribution of external shear values. 
\citet{HilbertEtal2009} performed similar ray-tracing experiments
through the Millennium Simulation \citep{SpringelEtal2005}, generating a
predicted $\kappa$ at every position in a simulated sky.
\citet{HilbertEtal2009} found that \MS lines of sight with strong lenses
were not biased towards high $\kappa$, although selection functions for
discovering or using particular samples of strong lenses were not taken
into account.

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappa$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappa)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuEtal2012b}.
% B1608$+$656 is indeed a high $\kappa$
% line of sight, but it does not reside in a massive cluster environment.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappa$ values to construct
$\Pr(\kappa|\data)$. For the most over-dense lines of sight the
Green et al. results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense lines.

In this paper, we combine the halo model reconstruction approach of
\citeauthor{GunnarssonEtal2006}, \citeauthor{JonssonEtal2010}, 
\citeauthor{WongEtal2011} and others, with the
idea of calibrating to simulations from \citeauthor{SuyuEtal2010}.
Taking the $\kappa$ values from
\citeauthor{HilbertEtal2009}'s ray-tracing calculations as the ``truth''
that we have to recover, we use the \MS mock galaxy catalogues first to
{\it calibrate} the reconstruction, and account for unseen mass and
voids, and then to {\it test} the accuracy of the line of sight mass
reconstruction under this assumption. Probabilistically assigning mass
and redshift to every observed galaxy in a given observed field, we
generate Monte Carlo sample line of sight mass distributions, and so
construct the PDF $\Pr(\kappa|\data)$ for that field. We then emulate
the combination of many such PDFs to quantify the residual biases that
would be translated to the global (including cosmological) parameters.
In doing so, we aim to answer the following questions:

@@

Throughout this paper magnitudes are given in the AB system and
we adopt the Millenium Simulation's ``concordance'' parameters for our reference cosmology, \ie
$h=0.73$, $\Omega_m=0.25$ and $\Omega_\Lambda=0.75$, where the symbols indicate
the Hubble Constant in units of 100 km s$^{-1}$ Mpc$^{-1}$ and the matter and
dark energy density of the Universe in units of the critical density.
.%\citep[e.g.\ ][]{KomEtal09}.

\begin{figure*}
\includegraphics[width=\textwidth]{figs/viewofalightcone.eps}
\caption[magcut]{Four different views of a slightly over-dense \MS line
of sight, with a $\kappa$ of 0.03. 
{\it Top row, left:} The positions of galaxies projected on the sky. The area
of the circles is proportional to observed $i$-band flux; the brightest
object shown has an $i$ magnitude of 17.7. 
{\it Top row, centre:} The angular sizes of halos projected on the sky.
Red and blue regions lie within the NFW scale radius and virial radius
of each halo, respectively. There are essentially no empty lightcones. 
{\it Top row, right:} The individual $\kappa$ contributions of
each halo, assuming a \citet{BMO} profile and the \citet{Neto2007}
mass-concentration relation. Comparison of
this panel and the centre panel reveals the relative importance of
proximity to the line of sight.
{\it Bottom:} A view along the redshift axis, showing only halos with
$|x|<0.3$ arcmin. The area of the points is proportional to each halo's
mass: the most massive halo shown has $1.6\times10^{12}\Msun$. The
optical axis is shown by the dashed line, while the dotted lines mark
the lens and source planes for a B1608-like strong lens.}
\label{fig:lightcone}
\end{figure*}

@@

\section{Theoretical Background}
\label{sec:theory}
\comments{
Convergence, or Ricci focusing, occurs when a gravitational lens focuses
the rays in a given bundle. This focusing can cause distant objects to
appear brighter, and larger than they would if the lens were removed.
The convergence from an isolated mass sheet is the ratio of the projected surface
mass density ($\Sigma$) \devided by the
critical surface mass density at the redshift of the mass sheet
($\Sigma_{\rm cr}$),
\be
\kappax= \frac {\Sigma}
                              {\Sigma_{\rm cr}(\zd,\zs)}
\ee
where 
\be 
\label{eq:sigcrit} 
\Sigma_{\rm cr}(\zd,\zs) \equiv \frac{c^2 D_{\rm os}}{4 \pi G D_{\rm od} D_{\rm ds}}.
\ee
and the $D$'s are the angular diameter distances between the objects
refered to in the subscripts: o is the observer, d is the deflector and
s is the source. If the surface mass density of the sheet exceeds
$\Sigma_{\rm cr}$, multiple images of the source can occur, otherwise
only one image will be observed, although that image will still be
perturbed relative to the unlensed case.

@@

If there is external convergence present that is not included in the
lens modeling, then the time delay distance -- inferred assuming $\kappax
= 0$ -- will be $(1-\kappax)$ more than the true value of $\Dt$:
\be 
\label{eq:MassSheet:Dtbias}
\Dt^{\rm{true}}=\frac{\Dt^{{\kappax = 0}}}{1-\kappax}.
\ee

We can hence estimate the true distance {\it only} if we have with additional
knowledge of $\kappax$. Since $\kappax$ is typically small the absolute
uncertainty on the estimate of $\kappax$ corresponds to the fractional
uncertainty with which time-delay distances can be \infered.

@@

In order to test the accuracy of our convergence estimates, we need to
know the true convergence for each line of sight. We cannot use  real
lines of sight for this,  but must use simulated lines of sight instead.
In this section we briefly review the Millenium Simulation, the ray
tracing calculations that have been carried out in it, and the mock
galaxy catalogs that have been produced.

The \MS \citep{SpringelEtal2005} is a cosmological N-body simulation of
Dark Matter structures in a cubic region approximately 680~Mpc in
comoving size, followed from redshift 127 to the present day. With an
approximate halo mass resolution of $2\times10^{10}{\rm M}_{\odot}$
(corresponding approximately to a galaxy with luminosity $0.1L^{*}$), it
provides a detailed prediction for the distribution of dark structures
present in the Universe, under the assumptions of the $\Lambda$CDM model
of hierarchical structure formation, cosmological densities as given at
the end of \Sref{sec:intro}, and $\sigma_8 = 0.9$.
Galaxy properties,
such as stellar masses, luminosities and colours, were assigned to the
simulated halos according to a semi-analytic model for galaxy formation
\citep{DeLucia+Blaizot2007}: the resulting predictions for the galaxy
luminosity function and correlation function match the observational
data very well.

@@

To define a test line of sight, we draw a random sky position from
within the convergence maps, bilinearly interpolate between the pixels
of the map, and store the result as the ``true'' convergence for that
line of sight, $\kappa_{\rm true}$.

\todo{Stefan}{Check this is all accurate and complete!}

@@

\subsection{Mock Photometric Galaxy Catalogs}
\label{sec:MS:mocks}

We construct mock photometric galaxy catalogs by selecting all \MS
galaxies within a circular aperture of a given angular radius centred on
the line of sight position, with apparent $i$-band magnitude brighter
than a given limit. The parent catalog, from \citet{HilbertEtal2011},
contains galaxy positions and magnitudes, all of which include lensing
effects (deflections and magnifications). We also have access to the
underlying galaxy halo masses and redshifts, which we will use to
calibrate the reconstruction, and also to explore any sources of bias
and scatter. Since the true convergence is that of only the dark matter,
we focus on reconstructing the dark matter halos alone. Our model for
doing this is outlined in the next section.

\todo{Stefan, Tom}{Check this is all accurate and complete!}

@@

We require a method for estimating the
convergence at any point on the sky, given a catalog of observed
galaxy properties such as positions, brightnesses, colours, and
possibly stellar masses and redshifts, for every galaxy in the field of
the lens, down to some magnitude limit. Since we cannot observe the
surface mass density of each halo directy, we need some way of
estimating the mass of each halo in the catalog, so that we can compute
its contribution to the total $\kappax$ along the line of sight \citep[as
in \eg][]{GunnarssonEtal2006,WongEtal2011,KarpenkaEtal2012}. 
This mass assignment recipe
will be uncertain, but can be calibrated with cosmological simulations, which
represent a significant additional source of information to what is present in
the data.

Transforming an observed photometometric galaxy catalogue into a 
catalogue of halo masses, positions, and redshifts, will enable us to
attempt a reconstruction of the convergence induced by every halo near a
line of sight. \new{We also need to account for the convergence due to dark structures and
the divergence due to voids; we do not model voids or dark structures, but their
effects are included statistically by calibration the the convergence in halos,
$\kappah$, against the ray-traced convergence, $\kappax$, for \MS lines of sight.}

@@

Cosmological dark matter simulations have shown that dark matter
halos are well-approximated by NFW profiles \citep{NFW}, with a density
profile given by
\be\label{eq:rhonfw}
\rho(r) = 
\frac{\rho_0}{\left(r/r_{s}\right)
\left(1+r/r_{s}\right)^{2}},
\ee
where $r_{s}$ is a characteristic radius of the cluster, representing where
the density slope transitions from $r^{-1}$ to $r^{-3}$. This radius is related
to the virial radius of the halo by $r_{s}~=~r_{200}/c$, where $r_{200}$ is the radius at 
which 
\be
\rho(r_{200}) = 200\rho_{{\mathrm{crit}}}(z), \text{ where } \rho_{\mathrm{crit}}(z) \equiv \frac{3H^{2}(z)}{8\pi G}
\ee
and $c$ is the concentration parameter, which can be estimated from the halo's mass,
using a mass--concentration relation. Typically more massive halos are less concentrated,
but there is some scatter; we use the relation of \citet{Neto2007} to estimate $c$
from the mass enclosed within $r_{200}$, which we denote as $M_{200}$. We find our results do not change if the \citet{MaccioEtal2008} Mass--concencetration relation is used instead. The best fit relation of \citet{Neto2007} is given by
\be
c_{200} = 4.67 (M_{200}/10^{14} h^{-1} \Msun)^{-0.11}.
\ee

@@

We use the relation of \citet{BehrooziEtal2010} to infer halo masses
from stellar masses; the details of this \proceedure are outlined in 
\Aref{appendix:MSMH}. Given an uncertain stellar mass and redshift for
each galaxy in the  catalog, we draw a sample halo mass from the PDF
that describes this uncertainty  (\Eref{eq:mhalo-mstar})  and use
\Eref{eq:kappai} to compute its contribution to the convergence,
$\kappa_i$; this can be done for all the halos along a specific line of sight
and summed to give a sample value of $\kappah$;  repeatedly applying
this \proceedure allows us to build up a histogram of $\kappah$ values
consistent with the data, and hence characterize $\pr(\kappah|\data)$.

@@

While halos contribute a positive convergence to $\kappa$, voids contribute
a negative convergence. When a ray bundle passes through a  region of space
that is less dense than $\rho_{\rm mean}(z)$ the rays are de-focused\footnote{In
principle, the convergence caused by halos is also caused by the overdensity,
but we neglect this effect since halos are typically much more dense than the
intergalactic medium}. Whilst the  full ray-traced solution takes into account
the effect of voids, our reconstruction cannot -- the halo catalogue does not
include voids. \new{In principle the absence of galaxies implies something about
the presence a void, but we do not attempt to use this information}. Neglecting voids
would lead to a heavily biased estimate of $\kappa$. Moreover, the halos in 
the model only add mass to the background density, and that background density
already accounts for all the mass in the Universe! In order to make voids
negative relative to the mean density, we must subtract a significant smooth
mass component. The problem is that we do not have a good model for the
density structure of that smooth component.

The solution we present here is to calibrate the relationship between halo
convergence $\kappah$ and underlying convergence $\kappa$ (as calculated
during full ray tracing) using the simulations themselves.
The halo modelling \proceedure outlined in \Sref{sec:model:halos} allows 
us to estimate 
the convergence due to halos given observations of 
mass, redshift and position,
$\pr(\kappah|\mathcal{D})$, but what we are interested in is the total convergence
along the line of sight $\pr(\kappa|\mathcal{D})$. We can obtain this by
considering the expression:
\begin{equation}
\Pr(\kappa|\mathcal{D}) = \int \dee\kappah 
   \Pr(\kappa|\kappah) \Pr(\kappah|\mathcal{D})
\label{eq:kappaconv}   
\end{equation}
The first term in the integrand relates the convergence due to model halos,
$\kappah$, to the true convergence, $\kappa$. This conditional distribution
can be constructed from the Millennium Simulation catalogs, by computing
$\kappah$ for each selected line of sight using the true values of the halo
parameters,  and accumulating $\kappah$ and $\kappa$ along a large number of
such lines of
sight to  characterize $\Pr(\kappa|\kappah)$.
For any given line of sight we can then estimate
$\pr(\kappah|\data)$ as described above, and then
multiply it by $\Pr(\kappa|\kappah)$ and integrate out the intermediate
nuisance parameter $\kappah$. 
% We can do the marginalisation over $\kappah$
% by drawing samples from the two-dimensional grid and keeping only the
% $\kappa$ values, to form our final result, $\Pr(\kappa|\mathcal{D})$.

However, given an uncertain stellar mass to halo mass conversion, the PDF
$\pr(\kappah|\{\Mstar,z\})$ can become skewed relative to
$\pr(\kappah|\{\Mhalo,z\})$; this is because of the assymetry in the
conversion from stellar mass through halo mass into $\kappah$. A long tail of
high halo masses gives rise to a tendency to overestimate $\kappah$.
If this shift is ignored, the resulting $\Pr(\kappa|\data)$ will be 
biased towards high values of $\kappa$.

>>>>

\title[Line of Sight Mass Reconstruction]
{Reconstructing the Lensing Mass in the Universe \\
from Photometric Catalogue Data}

\author[Collett \etal]{%
  Thomas~E.~Collett$^{1}$\thanks{\collettemail},
  Philip~J.~Marshall$^{2}$,
  Matthew~W.~Auger$^{1}$,
  Stefan~Hilbert$^{3}$,
\newauthor{%
  Sherry~H.~Suyu$^{4}$,
  Zachary~Greene$^{4}$,
  Tommaso~Treu$^{4}$\thanks{\packard},}
\newauthor{%
  Christopher~D.~Fassnacht$^{5}$,
  L\'eon~V.~E.~Koopmans$^{6}$,
  Roger~D.~Blandford$^{3}$} 
  \medskip\\
  $^1$\ioa\\
  $^2$\oxford\\
  $^3$\kipac\\
  $^4$\ucsb\\
  $^5$\davis\\
  $^6$\kapteyn
}

@@

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight.
Such massive structures contribute an ``external convergence'' $\kappax$ that
in time delay gravitational lenses can dominate the uncertainty in the inferred distances.  We
characterise this uncertainty by marginalising over a probability density
function for $\kappax$: here, we investigate the use of a simple halo model to
estimate $\pr(\kappax|\data)$ given noisy estimates of the photometric redshift
and stellar mass of galaxies observed in optical imaging surveys of the field
in question. We use mock catalogues from the Millennium Simulation, and compare
our reconstructed $\pr(\kappax|\data)$ to ray-traced $\kappax$ truth values: 
taking into account realistic uncertainties on redshift and stellar masses, we
typically find that the reconstruction yields a factor of 2 improvement in precision of
over the global $\pr(\kappax)$. For any given line of sight the
convergence estimate can be made before the investment of expensive follow-up
observations. We find that the lowest $\kappax$ lines of sight have the best constrained 
$\pr(\kappax)$ after applying our reconstruction. We find that selecting systems by the precision with which
$\kappax$ can be estimated leads to a sample of unbiased
time delay distance estimates with 1\% uncertainties due to external convergence effects.
Photometric data alone is sufficient to pre-select the best lenses and can be done
before investment in lightcurve monitoring. Preferentially selecting lines of
sight with high external shear (as might happen in an assembled sample of
quadruply-imaged time delay lenses) produces time-delay distance estimates with
typical uncertainties of 1.5\% and biases of 0.3\%. We suggest areas for the improvement of this general
analysis framework that should allow yet more accurate cosmological inferences
to be made.

@@

Large numerical simulations have been used to
predict, from first principles, the distribution of likely external
shears and convergences along various lines of sight, including those
containing strong lens systems.  \citet{Holder+Schechter2003} and
\citet{DalalEtal2005} carried out ray-tracing calculations in N-body
simulations to estimate the distribution of external shear values. 
\citet{HilbertEtal2009} performed similar ray-tracing experiments
through the Millennium Simulation \citep{SpringelEtal2005}, generating a
predicted $\kappax$ at every position in a simulated sky.
\citet{HilbertEtal2009} found that \MS lines of sight with strong lenses
were not biased towards high $\kappax$, although selection functions for
discovering or using particular samples of strong lenses were not taken
into account.

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappax$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappax)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuEtal2012b}.
% B1608$+$656 is indeed a high $\kappax$
% line of sight, but it does not reside in a massive cluster environment.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappax$ values to construct
$\Pr(\kappax|\data)$. For the most over-dense lines of sight the
Green et al. results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense lines.

In this paper, we combine the halo model reconstruction approach of
\citeauthor{GunnarssonEtal2006}, \citeauthor{JonssonEtal2010}, 
\citeauthor{WongEtal2011} and others, with the
idea of calibrating to simulations from \citeauthor{SuyuEtal2010}.
Taking the $\kappax$ values from
\citeauthor{HilbertEtal2009}'s ray-tracing calculations as the ``truth''
that we have to recover, we use the \MS mock galaxy catalogues first to
{\it calibrate} the reconstruction, and account for unseen mass and
voids, and then to {\it test} the accuracy of the line of sight mass
reconstruction under this assumption. Probabilistically assigning mass
and redshift to every observed galaxy in a given observed field, we
generate Monte Carlo sample line of sight mass distributions, and so
construct the PDF $\Pr(\kappax|\data)$ for that field. We then emulate
the combination of many such PDFs to quantify the residual biases that
would be translated to the global (including cosmological) parameters.
In doing so, we aim to answer the following questions:

@@

Throughout this paper magnitudes are given in the AB system and
we adopt the Millennium Simulation's ``concordance'' parameters for our reference cosmology, \ie
$h=0.73$, $\Omega_m=0.25$ and $\Omega_\Lambda=0.75$, where the symbols indicate
the Hubble Constant in units of 100 km s$^{-1}$ Mpc$^{-1}$ and the matter and
dark energy density of the Universe in units of the critical density.
.%\citep[e.g.\ ][]{KomEtal09}.

\begin{figure*}
\includegraphics[width=\textwidth]{figs/viewofalightcone.eps}
\caption[magcut]{Four different views of a slightly over-dense \MS line
of sight, with a $\kappax$ of 0.03. 
{\it Top row, left:} The positions of galaxies projected on the sky. The area
of the circles is proportional to observed $i$-band flux; the brightest
object shown has an $i$ magnitude of 17.7. 
{\it Top row, centre:} The angular sizes of halos projected on the sky.
Red and blue regions lie within the NFW scale radius and virial radius
of each halo, respectively. There are essentially no empty lightcones. 
{\it Top row, right:} The individual $\kappax$ contributions of
each halo, assuming a \citet{BMO} profile and the \citet{Neto2007}
mass-concentration relation. Comparison of
this panel and the centre panel reveals the relative importance of
proximity to the line of sight.
{\it Bottom:} A view along the redshift axis, showing only halos with
$|x|<0.3$ arcmin. The area of the points is proportional to each halo's
mass: the most massive halo shown has $1.6\times10^{12}\Msun$. The
optical axis is shown by the dashed line, while the dotted lines mark
the lens and source planes for a B1608-like strong lens.}
\label{fig:lightcone}
\end{figure*}

@@

\section{Theoretical Background}
\label{sec:theory}
\comments{
Convergence, or Ricci focusing, occurs when a gravitational lens focuses
the rays in a given bundle. This focusing can cause distant objects to
appear brighter, and larger than they would if the lens were removed.
The convergence from an isolated mass sheet is the ratio of the projected surface
mass density ($\Sigma$) \devided by the
critical surface mass density at the redshift of the mass sheet
($\Sigma_{\rm cr}$),
\be
\kappax= \frac {\Sigma}
                              {\Sigma_{\rm cr}(\zd,\zs)}
\ee
where 
\be 
\label{eq:sigcrit} 
\Sigma_{\rm cr}(\zd,\zs) \equiv \frac{c^2 D_{\rm os}}{4 \pi G D_{\rm od} D_{\rm ds}}.
\ee
and the $D$'s are the angular diameter distances between the objects
referred to in the subscripts: o is the observer, d is the deflector and
s is the source. If the surface mass density of the sheet exceeds
$\Sigma_{\rm cr}$, multiple images of the source can occur, otherwise
only one image will be observed, although that image will still be
perturbed relative to the unlensed case.

@@

If there is external convergence present that is not included in the
lens modelling, then the time delay distance -- inferred assuming $\kappax
= 0$ -- will be $(1-\kappax)$ more than the true value of $\Dt$:
\be 
\label{eq:MassSheet:Dtbias}
\Dt^{\rm{true}}=\frac{\Dt^{{\kappax = 0}}}{1-\kappax}.
\ee

We can hence estimate the true distance {\it only} if we have additional
knowledge of $\kappax$. Since $\kappax$ is typically small the absolute
uncertainty on the estimate of $\kappax$ corresponds to the fractional
uncertainty with which time-delay distances can be \infered.

@@

In order to test the accuracy of our convergence estimates, we need to
know the true convergence for each line of sight. We cannot use  real
lines of sight for this,  but must use simulated lines of sight instead.
In this section we briefly review the Millennium Simulation, the ray
tracing calculations that have been carried out in it, and the mock
galaxy catalogues that have been produced.

The \MS \citep{SpringelEtal2005} is a cosmological N-body simulation of
Dark Matter structures in a cubic region approximately 680~Mpc in
co-moving size, followed from redshift 127 to the present day. With an
approximate halo mass resolution of $2\times10^{10}{\rm M}_{\odot}$
(corresponding approximately to a galaxy with luminosity $0.1L^{*}$), it
provides a detailed prediction for the distribution of dark structures
present in the Universe, under the assumptions of the $\Lambda$CDM model
of hierarchical structure formation, cosmological densities as given at
the end of \Sref{sec:intro}, and $\sigma_8 = 0.9$.
Galaxy properties,
such as stellar masses, luminosities and colours, were assigned to the
simulated halos according to a semi-analytic model for galaxy formation
\citep{DeLucia+Blaizot2007}: the resulting predictions for the galaxy
luminosity function and correlation function match the observational
data very well.

@@

To define a test line of sight, we draw a random sky position from
within the convergence maps, bilinearly interpolate between the pixels
of the map, and store the result as the ``true'' convergence for that
line of sight, $\kappax^{\rm true}$.

@@

\subsection{Mock Photometric Galaxy Catalogues}
\label{sec:MS:mocks}

We construct mock photometric galaxy catalogues by selecting all \MS
galaxies within a circular aperture of a given angular radius centred on
the line of sight position, with apparent $i$-band magnitude brighter
than a given limit. The parent catalogue, from \citet{HilbertEtal2011},
contains galaxy positions and magnitudes, all of which include lensing
effects (deflections and magnifications). We also have access to the
underlying galaxy halo masses and redshifts, which we will use to
calibrate the reconstruction, and also to explore any sources of bias
and scatter. Since the true convergence is that of only the dark matter,
we focus on reconstructing the dark matter halos alone. Our model for
doing this is outlined in the next section.

@@

We require a method for estimating the
convergence at any point on the sky, given a catalogue of observed
galaxy properties such as positions, brightnesses, colours, and
possibly stellar masses and redshifts, for every galaxy in the field of
the lens, down to some magnitude limit. Since we cannot observe the
surface mass density of each halo directly, we need some way of
estimating the mass of each halo in the catalogue, so that we can compute
its contribution to the total $\kappax$ along the line of sight \citep[as
in \eg][]{GunnarssonEtal2006,WongEtal2011,KarpenkaEtal2012}. 
This mass assignment recipe
will be uncertain, but can be calibrated with cosmological simulations, which
represent a significant additional source of information to what is present in
the data.

Transforming an observed photometric galaxy catalogue into a 
catalogue of halo masses, positions, and redshifts, will enable us to
attempt a reconstruction of the convergence induced by every halo near a
line of sight. \new{We also need to account for the convergence due to dark structures and
the divergence due to voids; we do not model voids or dark structures, but their
effects are included statistically by calibration the the convergence in halos,
$\kappah$, against the ray-traced convergence, $\kappax$, for \MS lines of sight.}

@@

Cosmological dark matter simulations have shown that dark matter
halos are well-approximated by NFW profiles \citep{NFW}, with a density
profile given by
\be\label{eq:rhonfw}
\rho(r) = 
\frac{\rho_0}{\left(r/r_{s}\right)
\left(1+r/r_{s}\right)^{2}},
\ee
where $r_{s}$ is a characteristic radius of the cluster, representing where
the density slope transitions from $r^{-1}$ to $r^{-3}$. This radius is related
to the virial radius of the halo by $r_{s}~=~r_{200}/c$, where $r_{200}$ is the radius at 
which 
\be
\rho(r_{200}) = 200\rho_{{\mathrm{crit}}}(z), \text{ where } \rho_{\mathrm{crit}}(z) \equiv \frac{3H^{2}(z)}{8\pi G}
\ee
and $c$ is the concentration parameter, which can be estimated from the halo's mass,
using a mass--concentration relation. Typically more massive halos are less concentrated,
but there is some scatter; we use the relation of \citet{Neto2007} to estimate $c$
from the mass enclosed within $r_{200}$, which we denote as $M_{200}$. We find our results do not change if the \citet{MaccioEtal2008} Mass--concentration relation is used instead. The best fit relation of \citet{Neto2007} is given by
\be
c_{200} = 4.67 (M_{200}/10^{14} h^{-1} \Msun)^{-0.11}.
\ee

@@

We use the relation of \citet{BehrooziEtal2010} to infer halo masses
from stellar masses; the details of this \proceedure are outlined in 
\Aref{appendix:MSMH}. Given an uncertain stellar mass and redshift for
each galaxy in the  catalogue, we draw a sample halo mass from the PDF
that describes this uncertainty  (\Eref{eq:mhalo-mstar})  and use
\Eref{eq:kappai} to compute its contribution to the convergence,
$\kappa_i$; this can be done for all the halos along a specific line of sight
and summed to give a sample value of $\kappah$;  repeatedly applying
this \proceedure allows us to build up a histogram of $\kappah$ values
consistent with the data, and hence characterize $\pr(\kappah|\data)$.

@@

While halos contribute a positive convergence to $\kappax$, voids contribute
a negative convergence. When a ray bundle passes through a  region of space
that is less dense than $\rho_{\rm mean}(z)$ the rays are de-focused\footnote{In
principle, the convergence caused by halos is also caused by the overdensity,
but we neglect this effect since halos are typically much more dense than the
intergalactic medium}. Whilst the  full ray-traced solution takes into account
the effect of voids, our reconstruction cannot -- the halo catalogue does not
include voids. \new{In principle the absence of galaxies implies something about
the presence a void, but we do not attempt to use this information}. Neglecting voids
would lead to a heavily biased estimate of $\kappax$. Moreover, the halos in 
the model only add mass to the background density, and that background density
already accounts for all the mass in the Universe! In order to make voids
negative relative to the mean density, we must subtract a significant smooth
mass component. The problem is that we do not have a good model for the
density structure of that smooth component.

The solution we present here is to calibrate the relationship between halo
convergence $\kappah$ and underlying convergence $\kappax$ (as calculated
during full ray tracing) using the simulations themselves.
The halo modelling \proceedure outlined in \Sref{sec:model:halos} allows 
us to estimate 
the convergence due to halos given observations of 
mass, redshift and position,
$\pr(\kappah|\mathcal{D})$, but what we are interested in is the total convergence
along the line of sight $\pr(\kappax|\mathcal{D})$. We can obtain this by
considering the expression:
\begin{equation}
\Pr(\kappax|\mathcal{D}) = \int \dee\kappah 
   \Pr(\kappax|\kappah) \Pr(\kappah|\mathcal{D})
\label{eq:kappaconv}   
\end{equation}
The first term in the integrand relates the convergence due to model halos,
$\kappah$, to the true convergence, $\kappax$. This conditional distribution
can be constructed from the Millennium Simulation catalogues, by computing
$\kappah$ for each selected line of sight using the true values of the halo
parameters,  and accumulating $\kappah$ and $\kappax$ along a large number of
such lines of
sight to  characterize $\Pr(\kappax|\kappah)$.
For any given line of sight we can then estimate
$\pr(\kappah|\data)$ as described above, and then
multiply it by $\Pr(\kappax|\kappah)$ and integrate out the intermediate
nuisance parameter $\kappah$. 
% We can do the marginalisation over $\kappah$
% by drawing samples from the two-dimensional grid and keeping only the
% $\kappa$ values, to form our final result, $\Pr(\kappa|\mathcal{D})$.

However, given an uncertain stellar mass to halo mass conversion, the PDF
$\pr(\kappah|\{\Mstar,z\})$ can become skewed relative to
$\pr(\kappah|\{\Mhalo,z\})$; this is because of the asymmetry in the
conversion from stellar mass through halo mass into $\kappah$. A long tail of
high halo masses gives rise to a tendency to overestimate $\kappah$.
If this shift is ignored, the resulting $\Pr(\kappax|\data)$ will be 
biased towards high values of $\kappax$.

We find that we can correct this bias to some extent by artificially 
suppressing the tails of $\pr(\kappah|\{\Mstar,z\})$. Let the median of
this PDF be denoted $\kappah^{\rm med}$. We again take a large number of
calibration sightlines from the \MS catalogues, but this time generate a mock
stellar mass catalogue for each one. We then reconstruct each lightcone in the
same way as we would an observed field,
estimate $\pr(\kappah|\{\Mstar,z\})$, and extract its median. Accumulating
these samples and their corresponding true values of $\kappax$, we can form the 
conditional distribution $\pr(\kappax|\kappah^{\rm med})$. 
Then, when inferring $\kappax$ from an observed photometric catalogue, we would
estimate $\pr(\kappah|\{\Mstar,z\}_{\rm obs})$ 
using the halo model, compute the median value 
$\kappa_{\rm h,obs}^{\rm med}$, and then use the following alternative
version of 
\Eref{eq:kappaconv} to infer $\kappax$: 
\bea
\Pr(\kappax|\mathcal{D}) &=& \int \dee\kappah^{\rm med} 
   \Pr(\kappax|\kappah^{\rm med}) \Pr(\kappah^{\rm med}|\mathcal{D}) \notag \\
                        &=& \Pr(\kappax|\kappa_{\rm h,obs}^{\rm med}).
\label{eq:calkappaconv}   
\eea
Here, we have replaced $\Pr(\kappah^{\rm med}|\mathcal{D})$ with a delta
function centred at the observed median value $\kappa_{\rm h,obs}$. 
The procedure of reducing the full PDF for $\kappah$ to just its median is
risky: the simulated lightcones that have median $\kappah$ values to match are
not guaranteed to provide an accurate estimate of $\kappax$ via this PDF
look-up. We proceed with caution, quantifying the biases and PDF widths in the
sections below.
